### 2021-11-17 proposalPres.tex
Edited a figure so it takes up less empty space
Updated proposalPres.tex

% Copyright 2004 by Till Tantau <tantau@users.sourceforge.net>.
%
% In principle, this file can be redistributed and/or modified under
% the terms of the GNU Public License, version 2.
%
% However, this file is supposed to be a template to be modified
% for your own needs. For this reason, if you use this file as a
% template and not specifically distribute it as part of a another
% package/program, I grant the extra permission to freely copy and
% modify this file as you see fit and even to delete this copyright
% notice.

% \UseRawInputEncoding
\documentclass{beamer}

% There are many different themes available for Beamer. A comprehensive
% list with examples is given here:
% http://deic.uab.es/~iblanes/beamer_gallery/index_by_theme.html
% You can uncomment the themes below if you would like to use a different
% one:
%\usetheme{AnnArbor}
%\usetheme{Antibes}
%\usetheme{Bergen}
%\usetheme{Berkeley}
%\usetheme{Berlin}
%\usetheme{Boadilla}
%\usetheme{boxes}
%\usetheme{CambridgeUS}
%\usetheme{Copenhagen}
%\usetheme{Darmstadt}
%\usetheme{default}
%\usetheme{Frankfurt}
%\usetheme{Goettingen}
%\usetheme{Hannover}
%\usetheme{Ilmenau}
%\usetheme{JuanLesPins}
%\usetheme{Luebeck}
\usetheme{Madrid}
%\usetheme{Malmoe}
%\usetheme{Marburg}
%\usetheme{Montpellier}
%\usetheme{PaloAlto}
%\usetheme{Pittsburgh}
%\usetheme{Rochester}
%\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Warsaw}

\usepackage{pgfgantt}
\usepackage{todonotes}
\usepackage{media9}
\usepackage{fontawesome5}
\usepackage{subfigure}
\usepackage{booktabs,array}
\usepackage{tabulary}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage{arydshln}

\usepackage[ruled, vlined, linesnumbered]{algorithm2e} % For algorithms

\usepackage{amsmath} % For typesetting math

% Customize Warsaw color 
\setbeamercolor*{palette primary}{use=structure,fg=white,bg=red!50!black}
\setbeamercolor*{palette secondary}{use=structure,fg=white,bg=red!60!black}
\setbeamercolor*{palette tertiary}{use=structure,fg=white,bg=red!70!black}

% Customize Warsaw block title and background colors
\setbeamercolor{block title}{bg=red!50!black,fg=white}

\setbeamertemplate{bibliography item}{\insertbiblabel}  % insert bibliography numbers instead of symbol
\setbeamertemplate{caption}[numbered] % adds the figure or table number to the caption.

\title[HIL Plant Modeling]{Hardware-in-the-Loop Plant Modeling for Autonomous
  Vehicle}

% % A subtitle is optional and this may be deleted
\subtitle{Modeling, Simulation, and Testing}

\author[H.~Grady, N.~Nauman]{Hannah~Grady \and Nicholas~Nauman 
\linebreak Advisor:~Dr.~Suruz~Miah}
% - Give the names in the same order as the appear in the paper.
% - Use the \inst{?} command only if the authors have different
%   affiliation.

\institute[Bradley University] % (optional, but mostly needed)
{
  Department of Electrical and Computer Engineering\\
  Bradley University\\
  1501 W. Bradley Avenue\\
  Peoria, IL, 61625, USA
}
% - Use the \inst command only if there are several affiliations.
% - Keep it simple, no one is interested in your street address.

\date[October~7,~2021]{Thursday, October~7,~2021}

% - Either use conference name or its abbreviation.
% - Not really informative to the audience, more for people (including
%   yourself) who are reading the slides online

\logo{\hfill\href{http://www.bradley.edu}{\includegraphics[width=0.75cm]{figs/logoBU1-Print}}}  % place logo in every page

% \subject{Mobile Robot Localization}
% This is only inserted into the PDF information catalog. Can be left
% out.

% If you have a file called "university-logo-filename.xxx", where xxx
% is a graphic format that can be processed by latex or pdflatex,
% resp., then you can add a logo as follows:

% \pgfdeclareimage[height=0.5cm]{university-logo}{university-logo-filename}
% \logo{\pgfuseimage{university-logo}}

% Delete this, if you do not want the table of contents to pop up at
% the beginning of each subsection:
\AtBeginSubsection[]
{
  \begin{frame}<beamer>{Outline}
    \tableofcontents[currentsection,currentsubsection]
  \end{frame}
}

% Delete this, if you do not want the table of contents to pop up at
% the beginning of each section:
\AtBeginSection[]
{
  \begin{frame}<beamer>{Outline}
    \tableofcontents[currentsection]
  \end{frame}
}

% Let's get started
\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Outline} 
  \tableofcontents%[pausesections]
  % You might wish to add the option [pausesections]
\end{frame}

% Section and subsections will appear in the presentation overview
% and table of contents.
\section{Introduction}

\begin{frame}{Introduction}{}
    \begin{block}{Applications of Autonomous Vehicles}
    	\begin{itemize}
    		\item Autonomous vehicles are being developed by many companies for commercial and personal use
    		\item Modeling the steering system for autonomous vehicles is a challenging task
     		\item Models for autonomous vehicle subsystems must be very accurate due to safety factors
		\end{itemize}
    \end{block}
        \begin{figure}
			\centering
			\begin{minipage}[t]{0.4\textwidth}
				\centering
				\includegraphics[height=1.5cm]{figs/img/autonomousVehiclesAStuff}
				\caption{AutonomouStuff Vehicle Fleet\textsuperscript{a}}
				\label{fig:fleet}
			\end{minipage}
			\begin{minipage}[t]{0.4\textwidth}
				\centering
				\includegraphics[height=1.5cm]{figs/img/autonomousVehiclesSteering}
				\caption{Autonomous Vehicle Steering System\textsuperscript{b}}
				\label{fig:steerSystem}
			\end{minipage}
        \end{figure}
    \begin{tiny}
		\textsuperscript{a}https://hexagonpositioning.com/pi-brands/autonomoustuff\\\textsuperscript{b}https://www.autonews.com/article/20181105/OEM10/181109921/delphi-s-pace-award-winning-e-steer-an-autonomous-vehicle-building-block
    \end{tiny}
\end{frame}

\begin{frame}{Problem Statement}
  \begin{block}{Problem Statement}
    \begin{large}
      Model vehicle subsystems for a Lexus platform.
 \begin{itemize}
          \item Steering Model
          \item Acceleration Model
          \item Brake Model
          \item Shift Model
          \item Speed Model
          \item Speed Control Model
        \end{itemize} 
    \end{large}
  \end{block}
  \pause
  \begin{block}{Proposed Solution}
    \begin{large}
      A potential solution is to develop conventional System Identification techniques for modeling subsystems using available raw data.  
    \end{large}
  \end{block}
\end{frame}

%----------------------------------

\section{Literature Review}

\begin{frame}{Literature Review}
  \begin{block}{Existing Solutions}
 \begin{itemize}
        \item Use of System Identification Toolbox in MATLAB to create models from data~\cite{Adnan2010}
	 \begin{itemize}
		    \tiny
		    		\item Offers a variety of model choices
				\item Needs a large amount of raw data to produce an accurate model
	\end{itemize}
	\item Third order ARMAX model creates models using traditional methods of analysis~\cite{Li1999}
	 \begin{itemize}
		    \tiny
		    		\item Allows the use of traditional analysis methods to create a model 
				\item More room for error during calculations
	\end{itemize}
	\item Steer-By-Wire method as a system model~\cite{Saruchi2015}
	 \begin{itemize}
		    \tiny
		    		\item Can model systems with small non-linearities 
				\item Mechanical components replaced by electrical components 
	\end{itemize}
	\item Authors in~\cite{Hussain2011} illustrate identification of multiple-input single-output model for maximum power point tracking of photovoltaic system.  
	\begin{itemize}
	\tiny
		\item Fourth order ARX model ended up giving the best fit
	\end{itemize}
\end{itemize}
  \end{block}
    % \begin{figure}[b]
    %     \centering
    %     \includegraphics[width=0.45\textwidth]{figs/img/CompaRob}
    %     \caption{CompaRob}
    %     %\label{fig:sysBlockDiag}
    % \end{figure}
\end{frame}

%----------------------------------

\section{System Requirements}

\begin{frame}{System Requirements}
  \begin{block}{Specifications}
    The vehicle plant model will fulfill the requirements listed below:
\begin{itemize}
    \item The resulting plant model will consist of accurate subsystem models
    \item The subsystems can be used to create a HIL testbench
    \item The steering model can handle very small changes in steering angles
    \begin{itemize}
    		\item smooth out any non-linear discontinuities that would normally be measured by the steering motor, especially for small changes in the steering angle, depicted in \autoref{fig:nonlinGraph}.
    \end{itemize}
\end{itemize}
\begin{figure}[h]
    \centering
    \captionsetup{justification=centering, margin=3cm}
    \includegraphics[width=1in]{figs/inkscape/nonlinearBehavior}
    \caption{Steering Non-linear Behavior}
    \label{fig:nonlinGraph}
\end{figure}
  \end{block}
\end{frame}

%--------------------------------

\section{System Architecture}

\begin{frame}{System Architecture}
  \begin{block}{System Architecture}
    \begin{large}
      The overall system architecture of this project consists of six subsystems
      \begin{itemize}
          \item Steering Model
          \item Acceleration Model
          \item Brake Model
          \item Shift Model
          \item Speed Model
          \item Speed Control Model
        \end{itemize} 
    \end{large}
  \end{block}
\end{frame}

\begin{frame}{System Architecture}
\centering \includegraphics[width=.45\linewidth]{figs/inkscape/steeringModelArchitecture}\quad%
\centering \begin{minipage}[b][0.4\textheight][c]{.45\linewidth} \begin{enumerate} \item Steering System \item Brake System \item Acceleration System \end{enumerate} \end{minipage}\\[1em]
\centering \includegraphics[width=.45\linewidth]{figs/inkscape/brakeModelArchitecture}\quad%
\centering \includegraphics[width=.45\linewidth]{figs/inkscape/accelerationModelArchitecture}
\end{frame}

\begin{frame}{System Architecture}
\centering \includegraphics[width=.45\linewidth]{figs/inkscape/shiftModelArchitecture}\quad%
\centering \begin{minipage}[b][0.4\textheight][c]{.45\linewidth} \begin{enumerate} \item Shift System \item Speed Control System \item Speed System \end{enumerate} \end{minipage}\\[1em]
\centering \includegraphics[width=.45\linewidth]{figs/inkscape/speedControlModelArchitecture}\quad%
\centering \includegraphics[width=.45\linewidth]{figs/inkscape/speedModelArchitecture}
\end{frame}

\begin{frame}{System Architecture}
\begin{figure}[h]
    \centering
    \captionsetup{justification=centering, margin=3cm}
    \includegraphics[width=4in]{figs/inkscape/steeringModelBlockDiagram}
    \caption{Steering System Block Diagram}
    \label{fig:steerBlockDiag}
\end{figure}
\end{frame}

\begin{frame}{System Inputs and Outputs}
\begin{minipage}[t]{0.48\linewidth}
\tiny
	\begin{block}{System Inputs}
		\begin{itemize}
			\tiny
		    \item Steering Model
		    \begin{itemize}
		    \tiny
		    		\item Torque Voltages
		    		\begin{itemize}
		    		\tiny
		    			\item Desired Steering Wheel Angle
		    			\item Actual Steering Wheel Angle
		    		\end{itemize}
		    \end{itemize}
		    \item Acceleration Model
		    \begin{itemize}
		    \tiny
		    		\item Acceleration Pedal Voltages
		    \end{itemize}
		    \item Brake Model
		    \begin{itemize}
		    \tiny
		    		\item Brake Pedal Pressure Voltages
		    		\item Brake Pedal Stroke Voltages 
		    		\item Brake Pedal On/Off Switch
		    \end{itemize}
		    \item Shift Model
		    \begin{itemize}
		    \tiny
		    		\item Desired Shift Gear
		    \end{itemize}
		    \item Speed Model
		    \begin{itemize}
		    \tiny
		    		\item Acceleration Pedal Position
		    		\item Brake Pedal Position
		    		\item Shifter Actual Gear
		    \end{itemize}
		    \item Speed Control Model
		    \begin{itemize}
		    \tiny
		    		\item Desired Vehicle Speed
		    \end{itemize}
		\end{itemize}
	\end{block}
\end{minipage}%
\hfill%
\begin{minipage}[t]{0.48\linewidth}
\begin{block}{System Outputs}
		\begin{itemize}
			\tiny
		    \item Steering Model
		    \begin{itemize}
		    \tiny
		    		\item Wheel Angle
		    		\item Actual Steering Angle
		    \end{itemize}
		    \item Acceleration Model
		    \begin{itemize}
		    \tiny
		    		\item Acceleration Pedal Position
		    \end{itemize}
		    \item Brake Model
		    \begin{itemize}
		    \tiny
		    		\item Brake Pedal Position
		    		\item Brake Pressed
		    \end{itemize}
		    \item Shift Model
		    \begin{itemize}
		    \tiny
		    		\item Actual Shift Gear
		    \end{itemize}
		    \item Speed Model
		    \begin{itemize}
		    \tiny
		    		\item Vehicle Speed
		    \end{itemize}
		    \item Speed Control Model
		    \begin{itemize}
		    \tiny
		    		\item Vehicle Speed
		    \end{itemize}
		\end{itemize}
	\end{block}
\end{minipage}
\end{frame}

%----------------------------------

\section{System Identification}

\begin{frame}{System Identification}
  \begin{block}{System Identification}
 \begin{itemize}
        \item Developing mathematical models for a dynamic system
        \item Uses the measurement of the input and output signals
        \begin{itemize}
        		\item Single Input Single Output (SISO)
        		\item Single Input Multiple Output (SIMO)
        		\item Multiple Input Single Output (MISO)
        		\item Multiple Input Multiple Output (MIMO)
        \end{itemize}
        \item Achieve an accurate estimation of the system response for any given input
        \item Mathworks' MATLAB System Identification Toolbox
\end{itemize}
  \end{block}
\end{frame}

\subsection{Preliminary Work}

\begin{frame}{Preliminary Work}
  \begin{block}{Preliminary Work}
 \begin{itemize}
        \item Documentation on MATLAB's System Identification Toolbox and tutorials
        \item Literature Review 
        \item Data Collection for steering, braking and acceleration subsystems
\end{itemize}
  \end{block}
\end{frame}

%----------------------------------
\subsection{Experimental Setup}

\begin{frame}{Experimental Setup}
	\begin{block}{Control Method}
		\begin{itemize}
			\item Manual Mode
			\begin{itemize}
				\item Vehicle is not autonomous
				\item Torque voltages are sent by the ECU
			\end{itemize}
			\item By-Wire Mode
			\begin{itemize}
				\item Vehicle is autonomous
				\item Torque voltages are sent by the PACMod
				\item Torque voltages sent from the ECU are discarded by open-circuiting the motors
			\end{itemize}
		\end{itemize}
		\begin{figure}
			\centering
    			\captionsetup{justification=centering, margin=3cm}
    			\includegraphics[width=1in]{figs/img/picturesVisitToAStuff/dataColletionSetup1-20211007}
    			\caption{Autonomous Vehicle Data Collection Setup}
    			\label{fig:vehicleSetup}
		\end{figure}
	\end{block}
\end{frame}

%----------------------------------
\subsection{Parts List}

\begin{frame}{Parts List}
  \begin{block}{Software and Hardware}
 \begin{itemize}
        \item Software
        \begin{itemize}
        \small
        \item MATLAB's System Identification Toolbox
        \item Vector CANAlyzer
        \end{itemize}
	\item Hardware
	\begin{itemize}
	\small
	\item Laptop
	\item PAC Mod
	\item CAN Case 
	\item CAN Bus
	\end{itemize}
\end{itemize}
  \end{block}
\end{frame}

%----------------------------------
\subsection{Modeling}

\begin{frame}{Model Equations}
	\begin{block}{}
		\begin{itemize}
			\item Transfer Function Model: 
			\begin{equation}
				\frac{Y(s)}{U_A(s)} = \frac{b_{n}s^{n} + b_{n-1}s^{n-1} + ... + b_1s + b_0}{a_{m}s^{m} + a_{m-1}s^{m-1} + ... + a_1s + a_0}
			\end{equation}
			where m = number of poles and n = number of zeros
			\item ARX Model: 
			\begin{equation}
				A(z)y(t) = B(z)u(t) + e(t)
			\end{equation}
		\end{itemize}
	\end{block}
\end{frame}

%----------------------------------
\section{Validation and Testing}

\begin{frame}{Simulation Results}
  \begin{block}{Steering System}
 \begin{itemize}
	\item Best fit percentage of the by-wire steering estimated model is 85.54
	\item Best fit percentage of the manual steering estimated model is 90.27
 \end{itemize}
 \begin{figure}
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/img/byWireSteeringTransferFunctionModel} % first figure itself
        \caption{Output of Estimated By-Wire System Model}
        \label{fig:byWireSteerModel}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/img/manualSteeringTransferFunctionModel} % second figure itself
        \caption{Output of Estimated Manual System Model}
        \label{fig:manualSteerModel}
    \end{minipage}
\end{figure}
  \end{block}
\end{frame}

\begin{frame}{By-Wire Steering System Transfer Function}
\begin{minipage}[t]{0.48\linewidth}
	\begin{block}{Torque A}
		\tiny
		\begin{tabular}{ cc }   % top level tables, with 2 columns
			% leftmost table of the top level table
			\begin{tabular}{ |c|c| } 
				\hline
				b_0 & 3.44E-16 \\
				b_1 & -2.203E-15 \\
				b_2 & 1.24E-13 \\
				b_3 & 5.975E-13 \\
				b_4 & 3.001E-12 \\
				b_5 & 5.23E-12 \\
				\hline
			\end{tabular} &  % starting rightmost sub table
			% table 2
			\begin{tabular}{ |c|c| } 
				\hline
				a_0 & 1.949E-16 \\
				a_1 & 1.455E-14 \\
				a_2 & 8.797E-13 \\
				a_3 & 2.304E-11 \\
				a_4 & 9.041E-10 \\
				a_5 & 9.01E-9 \\
				a_6 & 2.714E-7 \\
				a_7 & 9.241E-7 \\
				a_8 & 1.946E-5 \\
				a_9 & 3.889E-5 \\
				a_{10} & 0.0005721 \\
				a_{11} & 0.0008035 \\
				a_{12} & 0.008695 \\
				a_{13} & 0.008922 \\
				a_{14} & 0.07499 \\
				a_{15} & 0.05422 \\
				a_{16} & 0.3631 \\
				a_{17} & 0.1693 \\
				a_{18} & 0.9394 \\
				a_{19} & 0.2118 \\
				a_{20} & 1 \\
				\hline
			\end{tabular} \\
		\end{tabular}
	\end{block}
\end{minipage}%
\hfill%
\begin{minipage}[t]{0.48\linewidth}
	\begin{block}{Torque B}
		\tiny
		\begin{tabular}{ cc }   % top level tables, with 2 columns
			% leftmost table of the top level table
			\begin{tabular}{ |c|c| } 
				\hline
				b_0 & -6.645E-15 \\
				b_1 & 2.161E-14 \\
				b_2 & -2.605E-14 \\
				b_3 & -4.744E-13 \\
				b_4 & -1.813E-13 \\
				b_5 & -4.858E-12 \\
				\hline
			\end{tabular} &  % starting rightmost sub table
			% table 2
			\begin{tabular}{ |c|c| } 
				\hline
				a_0 & 2.92E-15 \\
				a_1 & 2.758E-13 \\
				a_2 & 1.761E-11 \\
				a_3 & 4.228E-10 \\
				a_4 & 1.422E-8 \\
				a_5 & 1.561E-7 \\
				a_6 & 1.501E-6 \\
				a_7 & 8.426E-6 \\
				a_8 & 5.981E-5 \\
				a_9 & 0.0002016 \\
				a_{10} & 0.001226 \\
				a_{11} & 0.002649 \\
				a_{12} & 0.01456 \\
				a_{13} & 0.02044 \\
				a_{14} & 0.104 \\
				a_{15} & 0.09228 \\
				a_{16} & 0.4417 \\
				a_{17} & 0.2257 \\
				a_{18} & 1.025 \\
				a_{19} & 0.2305 \\
				a_{20} & 1 \\
				\hline
			\end{tabular} \\
		\end{tabular}
	\end{block}
\end{minipage}
\end{frame}

\begin{frame}{Manual Steering System Transfer Function}
\begin{minipage}[t]{0.48\linewidth}
	\begin{block}{Torque A}
		\tiny
		\begin{tabular}{ cc }   % top level tables, with 2 columns
			% leftmost table of the top level table
			\begin{tabular}{ |c|c| } 
				\hline
				b_0 & 6.516E4 \\
				b_1 & -2.944E5 \\
				b_2 & 2.328E5 \\
				b_3 & -2.487E5 \\
				b_4 & 8.25E4 \\
				b_5 & -3.038E4 \\
				\hline
			\end{tabular} &  % starting rightmost sub table
			% table 2
			\begin{tabular}{ |c|c| } 
				\hline
				a_0 & 1.224E5 \\
				a_1 & 4.227E5 \\
				a_2 & 2.202E6 \\
				a_3 & 3.54E6 \\
				a_4 & 7.98E6 \\
				a_5 & 7.62E5 \\
				a_6 & 1.071E7 \\
				a_7 & 6.745E6 \\
				a_8 & 6.85E6 \\
				a_9 & 3.006E6 \\
				a_{10} & 2.365E6 \\
				a_{11} & 7.37E5 \\
				a_{12} & 4.669E5 \\
				a_{13} & 1.021E5 \\
				a_{14} & 5.319E4 \\
				a_{15} & 7752 \\
				a_{16} & 3353 \\
				a_{17} & 287.5 \\
				a_{18} & 102.4 \\
				a_{19} & 3.653 \\
				a_{20} & 1 \\
				\hline
			\end{tabular} \\
		\end{tabular}
	\end{block}
\end{minipage}%
\hfill%
\begin{minipage}[t]{0.48\linewidth}
	\begin{block}{Torque B}
		\tiny
		\begin{tabular}{ cc }   % top level tables, with 2 columns
			% leftmost table of the top level table
			\begin{tabular}{ |c|c| } 
				\hline
				b_0 & 1.209E4 \\
				b_1 & 1.855E5 \\
				b_2 & -2.801E5 \\
				b_3 & 4.72E5 \\
				b_4 & -3.976E4 \\
				b_5 & 7.693E4 \\
				\hline
			\end{tabular} &  % starting rightmost sub table
			% table 2
			\begin{tabular}{ |c|c| } 
				\hline
				a_0 & 4.687E4 \\
				a_1 & 7.037E5 \\
				a_2 & 2.624E6 \\
				a_3 & 6.133E6 \\
				a_4 & 1.197E7 \\
				a_5 & 1.487E7 \\
				a_6 & 1.887E7 \\
				a_7 & 1.446E7 \\
				a_8 & 1.314E7 \\
				a_9 & 7.01E6 \\
				a_{10} & 4.79E6 \\
				a_{11} & 1.886E6 \\
				a_{12} & 9.836E5 \\
				a_{13} & 2.937E5 \\
				a_{14} & 1.149E5 \\
				a_{15} & 2.617E4 \\
				a_{16} & 7205 \\
				a_{17} & 1232 \\
				a_{18} & 200 \\
				a_{19} & 23.54 \\
				a_{20} & 1 \\
				\hline
			\end{tabular} \\
		\end{tabular}
	\end{block}
\end{minipage}
\end{frame}

\begin{frame}{Simulation Results}
  \begin{block}{Acceleration System}
 \begin{itemize}
	\item Best fit percentage of the by-wire acceleration estimated model is 97.69
	\item Best fit percentage of the manual acceleration estimated model is 98.3
 \end{itemize}
 \begin{figure}
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/img/byWireAccelArxModel} % first figure itself
        \caption{Output of Estimated By-Wire System Model}
        \label{fig:byWireAccelModel}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/img/manualAccelTransferFunctionModel} % second figure itself
        \caption{Output of Estimated Manual System Model}
        \label{fig:manualAccelModel}
    \end{minipage}
\end{figure}
  \end{block}
\end{frame}

\begin{frame}{By-Wire Acceleration System Transfer Function}
	\begin{block}{Fourth-Order ARX Model}
		\begin{equation}
			A(z)y(t) = B(z)u(t) + e(t)
		\end{equation}
		%
		where, 
		%
		\begin{equation}
			A(z) = 1 - 1.018z^{-1} - 0.002901z^{-2} + 0.4631z^{-3} - 0.2038z^{-4}
		\end{equation}
		\begin{equation}
			B_1(z) = -0.0207 - 0.01912z^{-1} - 0.02159z^{-2} - 0.03307z^{-3}
		\end{equation}
		\begin{equation}
			B_2(z) = 0.02017 + 0.01833z^{-1} + 0.09461z^{-2} + 0.05683z^{-3}
		\end{equation}
	\end{block}
\end{frame}

\begin{frame}{Manual Acceleration System Transfer Function}
\begin{minipage}[t]{0.48\linewidth}
	\begin{block}{Torque A}
		\tiny
		\begin{tabular}{ cc }   % top level tables, with 2 columns
			% leftmost table of the top level table
			\begin{tabular}{ |c|c| } 
				\hline
				b_0 & -5.741E6 \\
				b_1 & -2.644E7 \\
				b_2 & -2.797E6 \\
				b_3 & -1.304E7 \\
				b_4 & -3.203E6 \\
				b_5 & -1.159E6 \\
				b_6 & -5.519E5 \\
				\hline
			\end{tabular} &  % starting rightmost sub table
			% table 2
			\begin{tabular}{ |c|c| } 
				\hline
				a_0 & 2.307E7 \\
				a_1 & 8.369E7 \\
				a_2 & 9.137E8 \\
				a_3 & 1.992E9 \\
				a_4 & 3.753E9 \\
				a_5 & 5.454E9 \\
				a_6 & 5.548E9 \\
				a_7 & 5.869E9 \\
				a_8 & 3.453E9 \\
				a_9 & 3.161E9 \\
				a_{10} & 1.54E9 \\
				a_{11} & 9.654E8 \\
				a_{12} & 3.556E8 \\
				a_{13} & 1.789E8 \\
				a_{14} & 5.082E7 \\
				a_{15} & 2.071E7 \\
				a_{16} & 4.569E6 \\
				a_{17} & 1.498E6 \\
				a_{18} & 2.553E5 \\
				a_{19} & 6.532E4 \\
				a_{20} & 8443 \\
				a_{21} & 1559 \\
				a_{22} & 147.3 \\
				a_{23} & 15.5 \\
				a_{24} & 1 \\
				\hline
			\end{tabular} \\
		\end{tabular}
	\end{block}
\end{minipage}%
\hfill%
\begin{minipage}[t]{0.48\linewidth}
	\begin{block}{Torque B}
		\tiny
		\begin{tabular}{ cc }   % top level tables, with 2 columns
			% leftmost table of the top level table
			\begin{tabular}{ |c|c| } 
				\hline
				b_0 & -4.159E6 \\
				b_1 & 1.851E6 \\
				b_2 & -4.457E6 \\
				b_3 & 9.043E5 \\
				b_4 & -7.45E5 \\
				b_5 & 1.865E5 \\
				b_6 & -1.545E4 \\
				\hline
			\end{tabular} &  % starting rightmost sub table
			% table 2
			\begin{tabular}{ |c|c| } 
				\hline
				a_0 & 5.224E6 \\
				a_1 & 4.205E7 \\
				a_2 & 1.429E8 \\
				a_3 & 3.233E8 \\
				a_4 & 5.444E8 \\
				a_5 & 7.326E8 \\
				a_6 & 8.247E8 \\
				a_7 & 7.402E8 \\
				a_8 & 6.301E8 \\
				a_9 & 3.99E8 \\
				a_{10} & 2.731E8 \\
				a_{11} & 1.264E8 \\
				a_{12} & 7.223E7 \\
				a_{13} & 2.48E7 \\
				a_{14} & 1.216E7 \\
				a_{15} & 3.084E6 \\
				a_{16} & 1.324E6 \\
				a_{17} & 2.422E5 \\
				a_{18} & 9.264E4 \\
				a_{19} & 1.159E4 \\
				a_{20} & 4003 \\
				a_{21} & 307.4 \\
				a_{22} & 96.87 \\
				a_{23} & 3.451 \\
				a_{24} & 1 \\
				\hline
			\end{tabular} \\
		\end{tabular}
	\end{block}
\end{minipage}
\end{frame}

%----------------------------------
\section{Timeline and Milestones}

\begin{frame}{Timeline and Milestones}
  \begin{block}{Milestones}
 \begin{itemize}
        \item Model steering subsystem 
	\item Test subsystem models with HIL system 
	\item Final report and presentation 
\end{itemize}
  \end{block}
\end{frame}

%----------------------------------

%\begin{frame}{Other Considerations}
%  \begin{block}{Factors}
%  	\begin{small}
%      \begin{itemize}
%        \item Public Health: Autonomous vehicles can make roads safer, which can prevent accidents
%and driving fatalities and injuries. As a result, people will be able to live longer and
%healthier lives. By creating accurate models, we will help advance the development of
%autonomous vehicles and make this a reality.
%		\item Public Safety: This is a relevant concern, because the real world implementations of inaccurate models could be disastrous since it is a self-driving vehicle. We will address this by creating a rigorous test plan for our models.
%		\item Public Welfare: This factor is somewhat related to public safety and health in that we need to make sure that these vehicles are safe for people to use. How we will help
%ensure this in our project is the same as how we will ensure public safety, by making sure we throughly test our models.
%		\item Global Factors: 
%		\item Cultural Factors: 
%		\item Social Factors: 
%		\item Environmental Factors: 
%		\item Business Factors: 
%		\item Economic Factors: 
%      \end{itemize}
%     \end{small}
%  \end{block}
%\end{frame}

%----------------------------------

\section{Concluding Remarks}
\begin{frame}{Concluding Remarks}
  \begin{block}{Project goals}
    \begin{LARGE}
      \begin{itemize}
        \item Develop models of an autonomous vehicle's subsystems
        \begin{itemize}
          \item Steering Model
          \item Acceleration Model
          \item Brake Model
          \item Shift Model
          \item Speed Model
          \item Speed Control Model
        \end{itemize}
      \end{itemize}
    \end{LARGE}
  \end{block}
  \begin{block}{Anticipated Challenges}
    \begin{itemize}
      \item Achieving an acceptable accuracy and reliability for each model due to nonlinearities over small changes
    \end{itemize}
  \end{block}
\end{frame}

%----------------------------------

\section{References}

%\begin{frame}{References}
%  \bibliographystyle{IEEEtran}
%   \begin{itemize}
%     \item N. Rawashdeh, R. Haddad, O. Jadallah, and A. To’ma, “A person-following
% robotic cart controlled via a smartphone application: design and evaluation,”
% 09 2017, pp. 1–5.
%     \item M. M. Islam, A. Lam, H. Fukuda, Y. Kobayashi, and Y. Kuno, “An intelligent
% shopping support robot: understanding shopping behavior from 2d skeleton data using gru network,” ROBOMECH Journal, vol. 6, no. 1, 2019.
%     \item J. Sales, J. Marti, R. Marin Prades, E. Cervera, and P. Sanz, “Comparob: The
% shopping cart assistance robot,” International Journal of Distributed Sensor Networks, vol. 2016, pp. 1–15, 02 2016.
%     \item M. S. Miah, J. Knoll, and K. Hevrdejs, “Intelligent range-only mapping and navigation for mobile robots,” IEEE Transactions on Industrial Informatics, vol. 14, no. 3, pp. 1164–1174, 2018.
%     \item D. Li and S. Lane, “A novel and versatile parabolic reflector that significantly
% improves wi-fi reception at different distances and angles,” 2013.

%   \end{itemize}

% \end{frame}

%----------------------------------

\begin{frame}{References}

  \bibliographystyle{IEEEtran}
  \bibliography{bib/references.bib}

%\begin{itemize}
%\item
%@INPROCEEDINGS{Donjaroennon2021,
%  author={Donjaroennon, Natthapon and Nuchkum, Suphatchakan and Leeton, Uthen},
%  booktitle={2021 9th International Electrical Engineering Congress (iEECON)}, 
%  title={Mathematical model construction of DC Motor by closed-loop system Identification technique Using Matlab/Simulink}, 
%  year={2021},
%  volume={},
%  number={},
%  pages={289-292},
%  doi={10.1109/iEECON51072.2021.9440305}}
%\end{itemize}
% 	\begin{itemize}
% 		\item T. Xie, H. Jiang, X. Zhao, and C. Zhang, “A wi-fi-based wireless indoor position sensing system with multipath interference mitigation,” Sep 2019. [Online].
% Available: https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6767237/
% 		\item A. M. Ladd, K. E. Bekris, A. Rudys, L. E. Kavraki, and D. S. Wallach, “Robotics-
% based location sensing using wireless ethernet,” Wireless Networks, vol. 11, no. 1-2, p. 189–204, 2005.
% 		\item M. Lindhe, K. Johansson, and A. Bicchi, “An experimental study of exploiting multipath fading for robot communications,” Robotics: Science and Systems III, 2007.
% 		\item M. Lindhe and K. Johansson, “Using robot mobility to exploit multipath fading,”Wireless Communications, IEEE, vol. 16, pp. 30 – 37, 03 2009.
% 	\end{itemize}

\end{frame}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: