### 2022-04-13 finalReportPres.tex
Updated Final Presentation

% Copyright 2004 by Till Tantau <tantau@users.sourceforge.net>.
%
% In principle, this file can be redistributed and/or modified under
% the terms of the GNU Public License, version 2.
%
% However, this file is supposed to be a template to be modified
% for your own needs. For this reason, if you use this file as a
% template and not specifically distribute it as part of a another
% package/program, I grant the extra permission to freely copy and
% modify this file as you see fit and even to delete this copyright
% notice.

% \UseRawInputEncoding
\documentclass{beamer}

% There are many different themes available for Beamer. A comprehensive
% list with examples is given here:
% http://deic.uab.es/~iblanes/beamer_gallery/index_by_theme.html
% You can uncomment the themes below if you would like to use a different
% one:
%\usetheme{AnnArbor}
%\usetheme{Antibes}
%\usetheme{Bergen}
%\usetheme{Berkeley}
%\usetheme{Berlin}
%\usetheme{Boadilla}
%\usetheme{boxes}
%\usetheme{CambridgeUS}
%\usetheme{Copenhagen}
%\usetheme{Darmstadt}
%\usetheme{default}
%\usetheme{Frankfurt}
%\usetheme{Goettingen}
%\usetheme{Hannover}
%\usetheme{Ilmenau}
%\usetheme{JuanLesPins}
%\usetheme{Luebeck}
\usetheme{Madrid}
%\usetheme{Malmoe}
%\usetheme{Marburg}
%\usetheme{Montpellier}
%\usetheme{PaloAlto}
%\usetheme{Pittsburgh}
%\usetheme{Rochester}
%\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Warsaw}

\usepackage{pgfgantt}
\usepackage{todonotes}
\usepackage{media9}
\usepackage{subfigure}
\usepackage{booktabs,array}
\usepackage[font=tiny,labelfont=bf]{caption}
\usepackage{tabulary}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage{arydshln}
\usepackage{rotating}
\usepackage{amsmath,amssymb,amsfonts,bm}

% Customize Warsaw color 
\setbeamercolor*{palette primary}{use=structure,fg=white,bg=red!50!black}
\setbeamercolor*{palette secondary}{use=structure,fg=white,bg=red!60!black}
\setbeamercolor*{palette tertiary}{use=structure,fg=white,bg=red!70!black}

% Customize Warsaw block title and background colors
\setbeamercolor{block title}{bg=red!50!black,fg=white}

\setbeamertemplate{navigation symbols}{} % Remove navigation symbols
\setbeamertemplate{bibliography item}{\insertbiblabel}  % insert bibliography numbers instead of symbol
\setbeamertemplate{caption}[numbered] % adds the figure or table number to the caption.

% Set up footer manually
\setbeamertemplate{footline}
{
  \leavevmode%
  \hbox{%
  \begin{beamercolorbox}[wd=.45\paperwidth,ht=2.25ex,dp=1ex,center]{author in head/foot}%
    \usebeamerfont{author in head/foot}\insertshortauthor\hspace*{1em}(\insertshortinstitute)
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.4\paperwidth,ht=2.25ex,dp=1ex,center]{title in head/foot}%
    \usebeamerfont{title in head/foot}\insertshorttitle
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.15\paperwidth,ht=2.25ex,dp=1ex,center]{date in head/foot}%
    \usebeamerfont{date in head/foot}\insertframenumber{} / \inserttotalframenumber
  \end{beamercolorbox}}%
  \vskip0pt%
}

%==============================================================================
%     TITLE
%==============================================================================
\title[HIL Plant Modeling]{Hardware-in-the-Loop Plant Modeling for Autonomous Vehicle}

\author[H.~Grady, N.~Nauman]{Hannah~Grady \and Nicholas~Nauman 
\linebreak Advisor:~Dr.~Suruz~Miah}
% - Give the names in the same order as the appear in the paper.
% - Use the \inst{?} command only if the authors have different
%   affiliation.

\institute[Bradley University] % (optional, but mostly needed)
{
  Department of Electrical and Computer Engineering\\
  Bradley University\\
  1501 W. Bradley Avenue\\
  Peoria, IL, 61625, USA
}
% - Use the \inst command only if there are several affiliations.
% - Keep it simple, no one is interested in your street address.

\date[April~29,~2022]{Tuesday, April~29,~2022}

% - Either use conference name or its abbreviation.
% - Not really informative to the audience, more for people (including
%   yourself) who are reading the slides online

\logo{\hfill\href{http://www.bradley.edu}{\includegraphics[width=0.75cm]{figs/logoBU1-Print}}}  % place logo in every page

\subject{Plant Modeling for Autonomous Vehicle}
% This is only inserted into the PDF information catalog. Can be left
% out.

% If you have a file called "university-logo-filename.xxx", where xxx
% is a graphic format that can be processed by latex or pdflatex,
% resp., then you can add a logo as follows:

% \pgfdeclareimage[height=0.5cm]{university-logo}{university-logo-filename}
% \logo{\pgfuseimage{university-logo}}

% Delete this, if you do not want the table of contents to pop up at
% the beginning of each subsection:
\AtBeginSection[]
{
  \begin{frame}<beamer>{Outline}
    \tableofcontents[currentsection,currentsubsection]
  \end{frame}
}

%==============================================================================
%==============================================================================
%     START OF SLIDES
%==============================================================================
%==============================================================================

% Let's get started
\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Outline} 
  \tableofcontents%[pausesections]
  % You might wish to add the option [pausesections]
\end{frame}

% Section and subsections will appear in the presentation overview
% and table of contents.
\section{Introduction}

\begin{frame}{Introduction}{}
	\begin{block}{Applications of Autonomous Vehicles}
    	\begin{itemize}
    		\item Autonomous vehicles are being developed by many companies for commercial and personal use
    		\item Modeling different subsystems (such as, steering, brake,
          acceleration, $\ldots$)  for autonomous vehicles is a challenging task
     		\item Models for autonomous vehicle subsystems must be very accurate due to safety factors
		\end{itemize}
    \end{block}
        \begin{figure}
			\centering
			\begin{minipage}[t]{0.4\textwidth}
				\centering
				\includegraphics[height=1.5cm]{figs/img/autonomousVehiclesAStuff}
				\caption{AutonomouStuff Vehicle Fleet\textsuperscript{a}}
				\label{fig:fleet}
			\end{minipage}
			\begin{minipage}[t]{0.4\textwidth}
				\centering
				\includegraphics[height=1.5cm]{figs/img/autonomousVehiclesSteering}
				\caption{Autonomous Vehicle Steering System\textsuperscript{b}}
				\label{fig:steerSystem}
			\end{minipage}
        \end{figure}
    \begin{tiny}
		\textsuperscript{a}https://hexagonpositioning.com/pi-brands/autonomoustuff\\\textsuperscript{b}https://www.autonews.com/article/20181105/OEM10/181109921/delphi-s-pace-award-winning-e-steer-an-autonomous-vehicle-building-block
    \end{tiny}
\end{frame}

\begin{frame}{Introduction}{Problem Statement}
  \begin{block}{Problem Statement}
    \begin{Large}
    		Model autonomous vehicle subsystems for a Lexus platform
    \end{Large}
  \end{block}
  \pause
  \begin{block}{Proposed Solution}
    \begin{Large}
    		Use neural network modeling techniques to develop plant models
    \end{Large}
  \end{block}
\end{frame}

%----------------------------------

%----------------------------------
\section{Literature Review}

\begin{frame}{Literature Review}{}
	\begin{block}{Existing Solutions}
 	\begin{itemize}
        \item Use of System Identification Toolbox in MATLAB to create models from data~\cite{Adnan2010}
	 \begin{itemize}
     \tiny
		    		\item Offers a variety of model choices
				\item Needs a large amount of raw data to produce an accurate model
	\end{itemize}
	\item Third order ARMAX model creates models using traditional methods of analysis~\cite{Li1999}
	 \begin{itemize}
     \tiny
		    		\item Allows the use of traditional analysis methods to create a model 
				\item More room for error during calculations
	\end{itemize}
	\item Steer-By-Wire method as a system model~\cite{Saruchi2015}
	 \begin{itemize}
     \tiny
		    		\item Can model systems with small non-linearities 
				\item Mechanical components replaced by electrical components 
	\end{itemize}
	\item Authors in~\cite{Hussain2011} illustrate identification of multiple-input single-output model for maximum power point tracking of photovoltaic system.  
	\begin{itemize}
    \tiny
		\item Fourth order ARX model ended up giving the best fit
	\end{itemize}
\end{itemize}
  \end{block}
\end{frame}

%---------------------------
\section{System Identification Preliminaries}

\begin{frame}{System Identification Preliminaries}{}
  \begin{block}{Preliminary Work}
 \begin{itemize}
        \item Documentation on MATLAB's System Identification Toolbox and tutorials
        \item Literature Review 
        \item Data Collection for steering, braking and acceleration subsystems
\end{itemize}
  \end{block}

   \begin {block}{MATLAB Example: Dealing with Multi-Variable Systems: Identification and Analysis}
  \begin{itemize}
  	\item  Learned how to create an iddata object from a given data set 
  	\item Created a state space model and compared it to the validation data 
  	\item Tried creating submodels for some of the channels in order to try and get a better fit 
  	\item Then created a MISO model and learned how to merge the two SISO models that were created earlier 
  	\end{itemize}
  	\end{block}
\end{frame}

%-------------------------
\section{System Architecture}

\begin{frame}{System Architecture}{}
	\begin{block}{System Architecture}
    \begin{large}
      The overall system architecture of this project consists of six subsystems: 
      \begin{itemize}
          \item Steering
          \item Acceleration
          \item Brake %Model
          \item Shift %Model
          \item Speed %Model
          \item Speed control %Model
        \end{itemize} 
    \end{large}
  \end{block}
\end{frame}

\begin{frame}{System Architecture}{Subsystems}
	\centering \includegraphics[width=.45\linewidth]{figs/inkscape/steeringModelArchitecture}\quad%
	\centering \begin{minipage}[b][0.4\textheight][c]{.45\linewidth} \begin{enumerate} \item Steering System 			\item Brake System \item Acceleration System \end{enumerate} \end{minipage}\\[1em]
	\centering \includegraphics[width=.45\linewidth]{figs/inkscape/brakeModelArchitecture}\quad%
	\centering \includegraphics[width=.45\linewidth]{figs/inkscape/accelerationModelArchitecture}
\end{frame}

\begin{frame}{System Architecture}{Subsystems}
	\centering \includegraphics[width=.45\linewidth]{figs/inkscape/shiftModelArchitecture}\quad%
	\centering \begin{minipage}[b][0.4\textheight][c]{.45\linewidth} \begin{enumerate} \item Shift System \item 		Speed Control System \item Speed System \end{enumerate} \end{minipage}\\[1em]
	\centering \includegraphics[width=.45\linewidth]{figs/inkscape/speedControlModelArchitecture}\quad%
	\centering \includegraphics[width=.45\linewidth]{figs/inkscape/speedModelArchitecture}
\end{frame}

%---------------------------
\section{System Modeling using Neural Networks}

\begin{frame}{Modeling using Neural Networks}{}
	\begin{block}{}
		\begin{figure}
			\centering
      \tikzset{%
        input neuron/.style={
          circle,
          fill=green!50,
          minimum size=0.7cm
        },
        neuron missing/.style={
          draw=none, 
          scale=1,
          fill=white,
          text height=0.01cm,
          execute at begin node=\color{black}$\vdots$
        },
      }

      \tikzset{%
        hidden neuron/.style={
          circle,
          fill=blue!50,
          minimum size=0.7cm
        },
        neuron missing/.style={
          draw=none, 
          scale=1,
          fill=white,
          text height=0.01cm,
          execute at begin node=\color{black}$\vdots$
        },
      }

      \tikzset{%
        output neuron/.style={
          circle,
          fill=red!50,
          minimum size=0.7cm
        },
        neuron missing/.style={
          draw=none, 
          scale=1,
          fill=white,
          text height=0.01cm,
          execute at begin node=\color{black}$\vdots$
        },
      }

      \begin{tikzpicture}[x=1.5cm, y=1.5cm]

        \foreach \m/\l [count=\y] in {1,2,missing,3}
        \node [input neuron/.try, neuron \m/.try] (input-\m) at (0,2.5-\y) {};

        \foreach \m [count=\y] in {1,2,3,missing,4}
        \node [hidden neuron/.try, neuron \m/.try ] (hidden-\m) at (2,2.5-\y*1.1) {};

        \foreach \m [count=\y] in {1,missing,2}
        \node [output neuron/.try, neuron \m/.try] (output-\m) at (4,2.5-\y*1.2) {};

        \foreach \l [count=\i] in {1,2,n}
        \draw [<-] (input-\i) -- ++(-1,0)
        node [above, midway] {$\mathbf{e}_{k,\l}$};

        \foreach \l [count=\i] in {1,2,3,\mbox{$\bar{n}$}}
        \node [above] at (hidden-\i.north) {$\sigma_{\l}$};

        \foreach \l [count=\i] in {1,m}
        \draw [->] (output-\i) -- ++(1,0)
        node [above, midway] {$\hat{u}_{a,k}^{(\l)}$};

        \foreach \i in {1,...,3}
        \foreach \j in {1,...,4}
        \draw [->] (input-\i) -- (hidden-\j);

        \foreach \i in {1,...,4}
        \foreach \j in {1,2}
        \draw [->] (hidden-\i) -- (output-\j);

        \foreach \l [count=\x from 0] in {Input, Hidden, Ouput}
        \node [align=center, above] at (\x*2,1.9) {\l \\ layer};

        \node [align=center, below] at (3,-2.5) {Weight \\ $\mathbf{W}_a$};
        \draw[->,ultra thick](3,-2.5) -- (3,1.6);

      \end{tikzpicture}
  \caption{Actor neural network structure for approximating control input.}
  \label{fig:nnActor}
		\end{figure}
	\end{block}
\end{frame}

\begin{frame}
Therefore, approximated actor signal can be written as:
\begin{align}
  \label{eq:approximatePolicyActorWeights}
    \hat{\mathbf{u}}_{a,k} = \mathbf{W}_a^T\bm{\sigma}(\mathbf{e}_k),
  \end{align}
%
where $\bm{\sigma}:\mathbb{R}^n\to\mathbb{R}^{\bar{n}}$ with $\bar{n} = n(n+1)/2$ and components of weight matrix ${\bf W}_a$ shown in Fig.~\ref{fig:nnActor} are given by %
%
\begin{align*}
  {\bf W}_a = 
  \begin{bmatrix}
    w_a^{1,1} &  w_a^{1,2} & \ldots &w_a^{1,m}\\
    w_a^{2,1} & w_a^{2,2}& \ldots &w_a^{1,m}\\
    \vdots & \ddots & \ldots &\vdots\\
    w_a^{\bar{n},1} & w_a^{\bar{n},2}& \ldots &w_a^{\bar{n},m}
  \end{bmatrix}.
\end{align*}
%

Now take $\eta$ data points and form the set of linear equations as: %
%
\begin{align*}
&  \bm{\sigma}^T(\mathbf{e}_k) \mathbf{W}_a  \approx     \left(\mathbf{u}_{a,k}^{[d]}\right)^T\\
 & \bm{\sigma}^T(\mathbf{e}_{k+1}) \mathbf{W}_a  \approx     \left(\mathbf{u}_{a,k+1}^{[d]}\right)^T\\
                                            &\vdots \\
 & \bm{\sigma}^T(\mathbf{e}_{k+\eta-1}) \mathbf{W}_a  \approx     \left(\mathbf{u}_{a,k+\eta-1}^{[d]}\right)^T  
\end{align*}
%

\end{frame}

\begin{frame}
  In compact form:
  \begin{align*}
    \bm{\Sigma}\mathbf{W}_a \approx  \mathbf{U}^{[d]},\qquad\text{with}
    % \hat{\mathbf{U}} \approx \mathbf{U}^{[d]}, 
  \end{align*}
  %
  $\mathbb{R}^{\eta\times\bar{n}}\ni \bm{\Sigma} = [\bm{\Sigma}_0,\bm{\Sigma}_1,\ldots,\bm{\Sigma}_{\eta -1}]^T$ with $\bm{\Sigma}_\kappa = \bm{\sigma}^T({\bf e}_{k+\kappa})$ and $\mathbb{R}^{\eta\times m}\ni {\bf U}^{[d]} = [{\bf U}_0,{\bf U}_1,\ldots, {\bf U}_{\eta-1}]^T$ with ${\bf U}_\kappa = \left({\bf u}_{a,\kappa}\right)^T$ for $\kappa = 0,1,\ldots,\eta-1.$

  Let $\bm{\Sigma}\mathbf{W}_a\equiv \hat{\mathbf{U}},$ sum-of-squared error, $\delta_a$ can be written as:
  \begin{align*}
    \delta_a = \frac{1}{2}\|\mathbf{U}^{[d]} - \hat{\mathbf{U}}\|^2
  \end{align*}

Actor weight ${\bf W}_a$ can then be found minimizing the sum--of--square error $\delta_a$ defined by %
%
\begin{align*}
  \delta_a = \frac{1}{2}\|{\bf U}^{[d]} - \bm{\Sigma} {\bf W}_a\|^2  = \frac{1}{2}\mathrm{Tr}\left[({\bf U}^{[d]} - \bm{\Sigma} {\bf W}_a)^T({\bf U}^{[d]} - \bm{\Sigma} {\bf W}_a)\right].
\end{align*}
%
Taking the derivative of $\delta_a$ with respect to ${\bf W}_a$ yields 
%
\begin{align}
  \frac{\partial\delta_a}{\partial\mathbf{W}_a} = - \bm{\Sigma}^T\mathbf{U}^{[d]} + \bm{\Sigma}^T \bm{\Sigma}\mathbf{W}_a
  \label{eq:actorWeightsGradient}
\end{align}
%
Setting $\frac{\partial\delta_a}{\partial\mathbf{W}_a}$ to zero yeilds
\begin{align*}
  {\bf W}_a = \left(\bm{\Sigma}^T\bm{\Sigma}\right)^T\bm{\Sigma}^T{\bf U}^{[d]}.
\end{align*}

\end{frame}

\begin{frame}
$\bullet$ Actor weight is updated at each policy iteration $r=0,1,\ldots$ using gradient descent approach as:

\begin{multline}
  \mathbf{W}_a^{(r+1)} = \mathbf{w}_a^{(r)} - \ell_a\frac{\partial\delta_a}{\partial\mathbf{W}_a}= \mathbf{W}_a^{(r)} - \ell_a\left(-\bm{\Sigma}^T\mathbf{U}^{[d]} + \bm{\Sigma}^T \bm{\Sigma}\mathbf{W}_a^{(r)}\right)\\
  = \mathbf{w}_a^{(r)}-\ell_a \bm{\Sigma}^T\left(\bm{\Sigma}\mathbf{W}_a^{(r)} - \mathbf{U}^{[d]}\right)
  \label{eq:actorWeightUpdate}
\end{multline}

\end{frame}

\begin{frame}{Modeling using Neural Networks}{}
	\begin{block}{}
 \begin{itemize}
        \item Used MATLAB's Neural Network Time Series App 
        \item All models generated using the Bayesian Regularization Algorithm 
        \item Models trained using collected log data 
\end{itemize}
  \end{block}
\end{frame}

%-------------------------
\section{Modeling Vehicle Subsystems}

\begin{frame}{Modeling Vehicle Subsystems}{System Requirements}
	\begin{block}{Specifications}
	\begin{minipage}[t]{0.62\linewidth}    
    Proposed vehicle subsystem models are expected to fulfill requirements listed below:
	\begin{itemize}
    		\item Resulting plant model will consist of reasonably accurate subsystem models
    		\item Subsystems can be used to create a HIL testbench
    		\item Steering model can handle very small changes in steering angles
    \begin{itemize}
		\item smooth out any non-linear discontinuities that would normally be measured by the steering motor, especially for small changes in the steering angle, depicted in \autoref{fig:nonlinGraph}.
    \end{itemize}
	\end{itemize}
	\end{minipage}
	\begin{minipage}[t]{0.34\linewidth}
	\begin{figure}[h]
		\centering
    		\captionsetup{justification=centering, margin=0.5cm}
    		\includegraphics[width=1.5in]{figs/inkscape/nonlinearBehavior}
    		\caption{Steering Non-linear Behavior}
    		\label{fig:nonlinGraph}
	\end{figure}
	\end{minipage}
\end{block}
\end{frame}
%-------------------------
\section{Validation and Testing}

\begin{frame}{Validation and Testing}{Experimental Setup}
  \begin{block}{Software and Hardware}
 \begin{itemize}
        \item Software
        \begin{itemize}
        \small
        \item MATLAB's System Identification Toolbox
        \item Vector CANAlyzer
        \end{itemize}
	\item Hardware
	\begin{itemize}
	\small
	\item Laptop
	\item PACMod ECU
	\item CANCase 
	\item CAN bus
	\end{itemize}
\end{itemize}
  \end{block}
\end{frame}

\begin{frame}{Validation and Testing}{Experimental Setup}
	\begin{block}{Control Method}
		\begin{itemize}
			\item Manual Mode
			\begin{itemize}
				\item Vehicle is not autonomous
				\item Torque voltages are sent by the vehicle ECU
			\end{itemize}
			\item By-Wire Mode
			\begin{itemize}
				\item Vehicle is autonomous
				\item Torque voltages are sent by the PACMod ECU
				\item Torque voltages sent from the vehicle ECU are discarded by open-circuiting the motors
			\end{itemize}
		\end{itemize}
		\begin{figure}
			\centering
    			\captionsetup{justification=centering, margin=3cm}
    			\includegraphics[width=1in]{figs/img/picturesVisitToAStuff/dataColletionSetup1-20211007}
    			\caption{Autonomous Vehicle Data Collection Setup}
    			\label{fig:vehicleSetup}
		\end{figure}
	\end{block}
\end{frame}

\begin{frame}{Validation and Testing}{Validation Setup}
	\begin{figure}
		\centering \includegraphics[width=.45\linewidth]{figs/img/picturesVisitToAStuff/hilBenchAstuff}\quad%
		\centering \begin{minipage}[b][0.4\textheight][c]{.45\linewidth} \begin{itemize} \item HIL Bench 						\item Measuring output voltages to confirm results \end{itemize} \end{minipage}\\[1em]
		\centering \includegraphics[angle=270,width=.45\linewidth]{figs/img/picturesVisitToAStuff/aStuffVisit2Nick}\quad%
		\centering \includegraphics[angle=270,width=.45\linewidth]{figs/img/picturesVisitToAStuff/hilMeasurementHannah}
	\end{figure}
\end{frame}

\begin{frame}{Transfer Function Modeling}{Steering System}
	\begin{block}{}
 \begin{itemize}
	\item Best fit percentage of the by-wire steering estimated model is 83.25
	\item Best fit percentage of the manual steering estimated model is 90.27
 \end{itemize}
 \begin{figure}
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.6\textwidth]{figs/img/byWireSteeringTransferFunctionModel} % first figure itself
        \caption{Output of Estimated By-Wire System Model}
        \label{fig:byWireSteerModel}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.6\textwidth]{figs/img/manualSteeringTransferFunctionModel} % second figure itself
        \caption{Output of Estimated Manual System Model}
        \label{fig:manualSteerModel}
    \end{minipage}
\end{figure}
  \end{block}
\end{frame}

\begin{frame}{Transfer Function Modeling}{Acceleration System}
	\begin{block}{}
 \begin{itemize}
	\item Best fit percentage of the by-wire acceleration estimated model is 97.75
	\item Best fit percentage of the manual acceleration estimated model is 98.3
 \end{itemize}
 \begin{figure}
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.6\textwidth]{figs/img/byWireAccelArxModel} % first figure itself
        \caption{Output of Estimated By-Wire System Model}
        \label{fig:byWireAccelModel}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.6\textwidth]{figs/img/manualAccelTransferFunctionModel} % second figure itself
        \caption{Output of Estimated Manual System Model}
        \label{fig:manualAccelModel}
    \end{minipage}
\end{figure}
  \end{block}
\end{frame}

\begin{frame}{Transfer Function Modeling}{Brake System}
\begin{block}{}
 \begin{itemize}
	\item Only manual models were considered 
	\item There were no transfer function models that could satisfy both the best fit and error requirements 
 \end{itemize}
% \begin{figure}
%    \centering
%    \begin{minipage}{0.45\textwidth}
%        \centering
%        \includegraphics[width=0.6\textwidth]{figs/img/byWireSteeringTransferFunctionModel} % first figure itself
%        \caption{Output of Estimated By-Wire System Model}
%        \label{fig:byWireSteerModel}
%    \end{minipage}\hfill
%    \begin{minipage}{0.45\textwidth}
%        \centering
%        \includegraphics[width=0.6\textwidth]{figs/img/manualSteeringTransferFunctionModel} % second figure itself
%        \caption{Output of Estimated Manual System Model}
%        \label{fig:manualSteerModel}
%    \end{minipage}
%\end{figure}
  \end{block}
\end{frame}

\begin{frame}{Neural Network Modeling}{Steering System}
	\begin{block}{}
  		\begin{figure}[H]
  			\centering \includegraphics[width=.45\linewidth , height=.37\textheight]{figs/img/steeringSimulinkBlock.jpg}\quad%
			\centering \begin{minipage}[b][0.4\textheight][c]{.45\linewidth}  \begin{itemize}
			\item Only a few samples are just outside of the accuracy requirements
			\item Transfer function model did not meet the accuracy requirements
			\end{itemize} \end{minipage}\\[1em]
			\centering \includegraphics[width=.45\linewidth , height=.37\textheight]{figs/img/steeringNeuralNetworkTrainedOutput.jpg}\quad%
			\centering \includegraphics[width=.45\linewidth , height=.37\textheight]{figs/img/steeringNeuralNetworkTrainedOutput2.jpg}
  		\end{figure}
	\end{block}
\end{frame}

\begin{frame}{Neural Network Modeling}{Acceleration System}
	\begin{block}{}
  		\begin{figure}[H]
  			\centering \includegraphics[width=.45\linewidth , height=.37\textheight]{figs/img/accelerationSimulinkBlock.jpg}\quad%
			\centering \begin{minipage}[b][0.4\textheight][c]{.45\linewidth}  \begin{itemize}
			\item All samples within 2\% of the actual output
			\item Transfer function model did not have the same accuracy bounds
			\item Fewer concerns about connections in Simulink
			\end{itemize} \end{minipage}\\[1em]
			\centering \includegraphics[width=.45\linewidth , height=.37\textheight]{figs/img/accelNeuralNetworkTrainedOutput.jpg}\quad%
			\centering \includegraphics[width=.45\linewidth , height=.37\textheight]{figs/img/accelNeuralNetworkTrainedOutput2.jpg}
  		\end{figure}
	\end{block}
\end{frame}

\begin{frame}{Neural Network Modeling}{Brake System}
	\begin{block}{}
\begin{figure}[H]
  			\centering \includegraphics[width=.48\linewidth , height=.37\textheight]{figs/img/newBrakeSimulinkBlock.jpg}\quad%
			\centering \begin{minipage}[b][0.4\textheight][c]{.45\linewidth}  \begin{itemize}
			\item With this model, the error is kept below 5\%
			\item Able to easily track model performance
			\item Provides better results than the transfer function model
			\end{itemize} \end{minipage}\\[1em]
			\centering \includegraphics[width=.45\linewidth , height=.37\textheight]{figs/img/brake_new_neuralNetworkFig.jpg}\quad%
			\centering \includegraphics[width=.45\linewidth , height=.37\textheight]{figs/img/brake_new_neuralNetworkFigLog2Test.jpg}
  		\end{figure}

  \end{block}
\end{frame}

%------------------------------------------------------------------------------
%     SECTION BREAK
%------------------------------------------------------------------------------

\section{Conclusions and Future Work}

\begin{frame}{Conclusions and Future Work}
  \begin{block}{Conclusions}
    \begin{itemize}
      \item Transfer Function to Neural Network models
      \item Neural Network modeling worked best 
    \end{itemize}
  \end{block}
  \begin{block}{Future Work}
    \begin{itemize}
      \item Test models using Hardware-in-the-Loop
      \item Create new vehicle controllers 
    \end{itemize}
  \end{block}
\end{frame}

%------------------------------------------------------------------------------
%     SECTION BREAK
%------------------------------------------------------------------------------

\section{References}

\begin{frame}{References}
  \bibliographystyle{IEEEtran}
  \bibliography{bib/references.bib}
\end{frame}

%==============================================================================
%==============================================================================
%     END OF SLIDES
%==============================================================================
%==============================================================================

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: